---
title: 'STAT 627: Final Project'
author: "Meera Patel & Lindsay Beyak"
date: "4/21/2022"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. Choose a data set: Glass Identification Data Set from UCI ML Repository:
https://archive.ics.uci.edu/ml/datasets/glass+identification

(a) Data Type: Multivariate
(b) Default Task: Classification
(c) Attribute Types: Real
(d) # Instances: 214
(e) # Attributes: 10


# Load libraries
```{r}
library(tidyverse)
```

# Load data and clean
```{r}
# load data (renamed .data to .csv)
glass <- read.csv("glass.csv", header = FALSE)
head(glass)

# check for missing values
is.na(glass) # no missing values

# rename columns
glass %>%
  rename(c(ID = V1, RI = V2, Na = V3, Mg = V4, Al = V5, Si = V6, K = V7,
           Ca = V8, Ba = V9, Fe = V10, Type = V11)) -> glass
head(glass)
```


2.
(a) 
1. ID = Id number of glass from 1 to 214
2. RI = Refractive Index
3. Na = Sodium (unit measurement: weight percent in corresponding oxide, as are attributes 4-10)
4. Mg = Magnesium
5. Al = Aluminum
6. Si = Silicon
7. K = Potassium
8. Ca = Calcium
9. Ba = Barium
10. Fe = Iron
11. Type = Type of glass (class attribute):
  1. building_windows_float_processed = float processed glass in building windows
  2. building_windows_non_float_processed = non-float processed glass in building windows
  3. vehicle_windows_float_processed = float processed glass in vehicle windows
  4. vehicle_windows_non_float_processed (none in the database) = non-float processed glass in vehicle windows
  5. containers = glass used for containers
  6. tableware = glass used for tableware
  7. headlamps = glass used for headlamps
  
Overall goal of analysis: The goal is to analyze the type of glass found at a crime scene in order to correctly distinguish if it is a type of "float processed" glass or not, with the least amount of statistical error.

The response variable is Type. The predictor variables are ID, RI, Na, Mg, Al, Si, K, Ca, Ba, and Fe.

The categorical variables are ID and Type. The quantitative variables are RI, Na, Mg, Al, Si, K, Ca, Ba, and Fe.

(b) We checked for missing values above. There were none, so no variables were removed. The sample size of our full data set remains at 214 observations.

(c)
```{r}

```

