---
title: 'STAT 627: Final Project'
author: "Meera Patel & Lindsay Beyak"
date: "4/21/2022"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. Choose a data set: Glass Identification Data Set from UCI ML Repository:
https://archive.ics.uci.edu/ml/datasets/glass+identification

(a) Data Type: Multivariate
(b) Default Task: Classification
(c) Attribute Types: Real
(d) # Instances: 214
(e) # Attributes: 10


# Load libraries
```{r}
library(tidyverse)
library(e1071)
library(ggplot2)
```

# Load data and clean
```{r}
# load data (renamed .data to .csv)
glass <- read.csv("glass.csv", header = FALSE)
head(glass)

# check for missing values
is.na(glass) # no missing values

# rename columns
glass %>%
  rename(c(ID = V1, RI = V2, Na = V3, Mg = V4, Al = V5, Si = V6, K = V7,
           Ca = V8, Ba = V9, Fe = V10, Type = V11)) -> glass
head(glass)
View(glass)
```


2.
(a) 
1. ID = Id number of glass from 1 to 214
2. RI = Refractive Index
3. Na = Sodium (unit measurement: weight percent in corresponding oxide, as are attributes 4-10)
4. Mg = Magnesium
5. Al = Aluminum
6. Si = Silicon
7. K = Potassium
8. Ca = Calcium
9. Ba = Barium
10. Fe = Iron
11. Type = Type of glass (class attribute):
  1. building_windows_float_processed = float processed glass in building windows
  2. building_windows_non_float_processed = non-float processed glass in building windows
  3. vehicle_windows_float_processed = float processed glass in vehicle windows
  4. vehicle_windows_non_float_processed (none in the database) = non-float processed glass in vehicle windows
  5. containers = glass used for containers
  6. tableware = glass used for tableware
  7. headlamps = glass used for headlamps
  
Overall goal of analysis: The goal is to analyze the type of glass found at a crime scene in order to correctly distinguish if it is a type of "float processed" glass or not, with the least amount of statistical error.

The response variable is Type. The predictor variables are ID, RI, Na, Mg, Al, Si, K, Ca, Ba, and Fe.

The categorical variables are ID and Type. The quantitative variables are RI, Na, Mg, Al, Si, K, Ca, Ba, and Fe.

(b) We checked for missing values above. There were none, so no variables were removed. The sample size of our full data set remains at 214 observations.

(c) Randomly create a validation/test data set that is approximately 10% of your full data set. Separate this out from your full data set. Treat the remaining 90% of your data set as your training data.
```{r}
# make new variable called "FloatProcessed" and Types 1,3 = Yes (1) and Types 2, 4, 5, 6, 7 = No (0)
head(glass)
glass %>%
  mutate(FloatProcessed = Type) -> glass

# convert 3 to 1 (1 is already 1)
glass$FloatProcessed[glass$FloatProcessed == 3] <- 1

# convert 2, 4, 5, 6, 7 to 0
glass$FloatProcessed[glass$FloatProcessed == 2] <- 0
glass$FloatProcessed[glass$FloatProcessed == 4] <- 0
glass$FloatProcessed[glass$FloatProcessed == 5] <- 0
glass$FloatProcessed[glass$FloatProcessed == 6] <- 0
glass$FloatProcessed[glass$FloatProcessed == 7] <- 0

# convert 1 to Yes and 0 to No
glass$FloatProcessed[glass$FloatProcessed == 1] <- "Yes"
glass$FloatProcessed[glass$FloatProcessed == 0] <- "No"

# convert Type and FloatProcessed to factor
glass$Type <- as.factor(glass$Type)
glass$FloatProcessed <- as.factor(glass$FloatProcessed)
View(glass)

# create new variable that takes average of Na, Mg, Al, Si, K, Ca, Ba, Fe
glass$OxideNum <- rowMeans(glass[ , c(3, 4, 5, 6, 7, 8, 9, 10)], na.rm = TRUE)
View(glass)

# check expected sample sizes of test/training data sets
214*.9 # 90% of training data = 192.6
214*.1 # 10% of test data = 21.4

set.seed(2200)
n <- length(glass$ID)
Z <- sample(n,n/10)
sort((1:n)[Z])
sort((1:n)[-Z])

train <- glass[sort(Z),] # 21 observations
test <- glass[-sort(Z),] # 193 observations

head(train)
View(train)
View(test)
```
Since the goal of our analysis is to distinguish whether or not the type of glass is float processed or non-float processed, we converted the float processed Types (1 and 3) to Yes and non-float processed/other glass Types (2, 4, 5, 6, 7) to No.

3.
(a) 
1. The first method is the Support Vector Classifier. 
  The model is: f(x*) = B0 + B1X1* + B2X2* + ... + BpXp*
  The tuning parameter is the radial kernel.
```{r}
# choose best kernel
set.seed(100)
svmt <- tune(svm, FloatProcessed ~ ., data = glass, ranges = list(kernel = c("linear", "polynomial", "radial", "sigmoid")))
svmt
summary(svmt)
```
  
2. The second method is Decision Trees.
  The model is
  The tuning parameter is max_features.

3. The third method is Clustering.
  The model is 

  The tuning parameter is K-means.

(b) Exploratory Data Analysis
```{r}
head(train)
# ggpairs(train)
```

(c)

(d)
1. Support Vector Classifier Analysis
```{r}
# tune
set.seed(100)
svmt <- tune(svm, FloatProcessed ~ ., data = glass, ranges = list(kernel = c("linear", "polynomial", "radial", "sigmoid")))
svmt
summary(svmt)

# tune on train data set
svm.tuned <- svm(FloatProcessed ~ RI + OxideNum, data = train, kernel = "radial")
svm.tuned
plot(svm.tuned, train, RI ~ OxideNum) # OxideNum is the average of the element variables
plot(svm.tuned, train, RI ~ Si) # Silicon is the main element in glass
yhat = predict(svm.tuned)
View(yhat)
table(yhat, train$FloatProcessed)
mean(yhat != train$FloatProcessed)

# ROC
pred <- prediction(as.numeric(yhat == "Yes"), as.numeric(train$FloatProcessed=="Yes"))
perf <- performance(pred, "tpr", "fpr")
perf # doesn't work because only tpr and fpr in confusion matrix
plot(perf, colorize = T, lwd = 2)
abline(a = 0, b = 1)
abline(h = 1)
abline(v = 0)
performance(pred, measure = "auc")@y.values

summary(glm(FloatProcessed ~ ., data = train, family = 'binomial'))

# tune on test data set
svm.tuned <- svm(FloatProcessed ~ RI + OxideNum, data = test, kernel = "radial")
svm.tuned
plot(svm.tuned, test, RI ~ OxideNum) # OxideNum is the average of the element variables
plot(svm.tuned, test, RI ~ Si) # Silicon is the main element in glass
yhat = predict(svm.tuned)
View(yhat)
table(yhat, test$FloatProcessed)
mean(yhat != test$FloatProcessed)

# ROC
pred <- prediction(as.numeric(yhat == "Yes"), as.numeric(test$FloatProcessed == "Yes"))
perf <- performance(pred, "tpr", "fpr")
perf
plot(perf, colorize = T, lwd = 2)
abline(a = 0, b = 1)
abline(h = 1)
abline(v = 0)
performance(pred, measure = "auc")@y.values

summary(glm(FloatProcessed ~ ., data = test, family = 'binomial'))
```

